---
title: "Food Dashboard"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
---
```{r setup, include=FALSE}
#multiple pages within one dashboard
#https://rmarkdown.rstudio.com/flexdashboard/layouts.html#multiple_pages

# Always start with a fresh environment
rm(list = ls())

source("Ryan_Code/data_cleaning.R")


library(flexdashboard)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(skimr)
library(randomForest)
library(caret)
library(e1071)
library(corrplot)
library(MASS)
library(gt)
library(Metrics)
library(gridExtra)
library(pscl)
library(boot)
library(rpart)
library(rpart.plot)
library(rgeos)
library(tigris)

cps_fexpend <- cps[!is.na(cps$fexpend),]

cps_fexpend = subset(cps_fexpend, select = -c(id, weight, fsecurity))

cps_fsecurity <- cps[!is.na(cps$fsecurity),]

cps_fsecurity = subset(cps_fsecurity, select = -c(id, weight, fexpend))


ggplot(data = cps_fsecurity, aes (x = fsecurity))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + labs(x = "Food Insecure", y = "Number of Households")

# ANALYSIS OF DISABILITY VARIABLE

cps_fsecurity$disability_cat = ifelse(cps_fsecurity$disability > 0, "Yes", "No")
cps_fsecurity$disability_cat = as.factor(cps_fsecurity$disability_cat)

cps_fsecurity$fsecurity_cat = ifelse(cps_fsecurity$fsecurity > 0, "yes", "no")
cps_fsecurity$fsecurity_cat = as.factor(cps_fsecurity$fsecurity_cat)
#NOTE: "-----------------------------------------------------------------------" is the section delimiter
#NOTE: "===================================== " is the PAGE delimiter

```
Data Overview{data-icon="fa-signal"}
===

Row {data-height=200}
-----------------------------------------------------------------------

### Goal of Research:
<font size = 2>
The goal of this analysis is to find determinants of food insecurity. The reason for doing so is humanitarian in nature as the
increase in efficacy of food distribution centers by determining higher risk areas should decrease the level of food. </font>


### Process:
<font size = 2>
In order to accomplish said goal we will begin by performing an exploratory analysis of all of the most important variables as determined by the variable importance plot. We will then proceed to demonstrate the effects of said variables on the response variables food security and food expenditure using regression analysis. From there we will use a random forest, which is a form of prediction via machine learning, to create predictions for both response variables for national block groups.
</font>


Row {data-height = 400}
---
### Key Variables
<font size = 2>
Variables used in order to predict levels of food insecurity and amount of USD used to buy food for one person in each household for a week are household size, number of elderly within household, number of hispanics within houshold, number of blacks within household, number of kids within household, number of individuals with an associates or higher within a household, number of employed within a household, number of married within a household, and number of disabled within a household. 
</font>

Exploratory Analysis - Food Security {data-navmenu="Exploratory Analysis"}
===

Column {.tabset}
---
### Disability Analysis 
```{r, fig.width = 20, fig.height = 12}
p1 <- ggplot(data = cps_fsecurity, aes(x = disability_cat, fill = disability_cat))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "Disabled Individual Living Within Household", y = "Number of Households", fill = "If Disabled") + scale_fill_brewer(palette = "Blues")

cps_disability <- cps_fsecurity %>% group_by(disability_cat) %>% summarise(Average = mean(fsecurity))

p2 <- ggplot(aes(x = disability_cat, y = Average, fill = Average), data = cps_disability ) + geom_bar(stat = "Identity") +
  labs(x = "Disabled Individual Living Within Household", y = "Average Level of Food Insecurity", fill = "Average Level") +
  scale_fill_distiller(palette = "Blues")

p3 <- ggplot() + geom_boxplot(aes(group = disability_cat, x = disability_cat, y = fsecurity, fill = disability_cat), data = cps_fsecurity) +
  labs(x = "Disabled Individual Living Within Household", y = "Level of Food Insecurity - log scale", fill = "If Disabled") + 
  scale_fill_brewer(palette = "Blues") + scale_y_log10() 

grid.arrange(p1, p2, p3, nrow = 1)
```

### Elderly Analysis
```{r, fig.width = 20, fig.height = 12}
g1 <- ggplot(data = cps_fsecurity, aes(x = elderly, fill = elderly))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "Number of Elderly in Household", y = "Number of Households") + scale_fill_fermenter(palette = "Blues") 

cps_elderly <- cps_fsecurity %>% group_by(elderly) %>% summarise(meld = mean(fsecurity))

g2 <- ggplot(aes(x = elderly, y = meld, fill = elderly), data = cps_elderly) + geom_bar(stat = "Identity") + 
  labs(x = "Number of Elderly in Household", y = "Average Level Of Food Insecurity", fill = "Number of Elderly") +
  scale_fill_distiller(palette = "Blues")

g3 <- ggplot() + geom_boxplot(aes(group = elderly, x = elderly, y = fsecurity, fill = elderly), data = cps_fsecurity) + 
  scale_y_log10() + scale_fill_distiller(palette = "Blues") + labs(x = "Number of Elderly Within Household", y = "Food Security Level Per Household - Log Scale", fill = "Number of Elderly")


grid.arrange(g1, g2, g3, nrow = 1)
```

### Education Analysis
```{r, fig.width = 20, fig.height = 12}
q1 <- ggplot(data = cps_fsecurity, aes(x = education))+ geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) +
  labs(x = "Number of Educated Individuals Within Household", y = "Number of Households")

cps_education <- cps_fsecurity %>% group_by(education) %>% summarise(med = mean(fsecurity))

q2 <- ggplot(aes(x = education, y = med, fill = education), data = cps_education) + geom_bar(stat = "Identity") +
  labs(x = "Number of Educated Individuals Within Household", y = "Average Level of Food Insecurity", fill = "Number of Educated") +
  scale_fill_distiller(palette = "Blues")

q3 <- ggplot() + geom_boxplot(aes(group = education, x = education, y = fsecurity, fill = education), data = cps_fsecurity) +
  scale_y_log10() + scale_fill_distiller(palette = "Blues") + labs(x = "Number of Individuals With Associates or Higher Within Household", y = "Food Security Level Per Household - log scale", fill = "Number of Educated")


grid.arrange(q1, q2, q3, nrow = 1)
```

### Employed Analysis
```{r, fig.width = 20, fig.height = 12}
p1 <- ggplot(data = cps_fsecurity, aes(x = employed))+ geom_bar() + geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of Employed Individuals Within Household", y = "Number of Households")

cps_employed <- cps_fsecurity %>% group_by(employed) %>% summarise(memp = mean(fsecurity))

p2 <- ggplot(aes(x = employed, y = memp, fill = employed), data = cps_employed) + geom_bar(stat = "Identity") +
  labs(x = "Number of Employed Individuals Within Household", y = "Average Level of Food Insecurity", fill = "Number of Employed") + scale_fill_distiller(palette = "Blues")
  
p3 <- ggplot() + geom_boxplot(aes(group = employed, x = employed, y = fsecurity, fill = employed), data = cps_fsecurity) +
  scale_y_log10() + scale_fill_distiller(palette = "Blues") + labs(x = "Number of Employed Individuals Within Household", y = "Food Security Level Per Household - log scale", fill = "Number of Employed")

grid.arrange(p1, p2, p3, nrow = 1)
```

### Household Size Analysis
```{r, fig.width = 20, fig.height = 12}
k1 <- ggplot(data = cps_fsecurity, aes(x = round(hhsize,0))) + geom_bar() + geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of Family Members Within Household", y = "Number of Households")

k2 <- ggplot(data = cps_fexpend, aes(x = fexpend))+geom_histogram(binwidth = 5)+ labs(x = "Food Expense", y = "Numer of Households" )

grid.arrange(k1, k2, nrow = 1)
```


Exploratory Analysis - Food Expenditure {data-navmenu="Exploratory Analysis"}
===
Column {.tabset}
---
```{r, warning=FALSE, show=FALSE}
cps_fexpend_f <- cps_fexpend

cps_fexpend_f$disability <- ifelse(cps_fexpend_f$disability > 0, "Yes", "No")

breaks <- c(0, 1, 2, 3, 4, 5, 6, 7, 8 ,9 ,10, 11, 12, 13)

tags <- c('0','1', '2', '3', '4', '5', '6', '7', '8' ,'9' ,'10', '11', '12')

cps_fexpend_f$hhsize_f <- cut(cps_fexpend_f$hhsize, breaks = breaks, include.lowest = TRUE, right = FALSE, labels = tags)

```


### Disability Analysis
```{r, fig.width = 20, fig.height = 12}
f1 <- ggplot(data = cps_fexpend_f, aes (x = disability, fill = disability))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "If Disabled Individual Within Household", y = "Number of Households", fill = "If Disabled") + scale_fill_brewer(palette = "BuGn")

cps_fexpend_disability <- cps_fexpend_f %>% group_by(disability) %>% summarise(me = mean(fexpend))

f2 <- ggplot(aes(x = disability, y = me, fill = me), data = cps_fexpend_disability) + geom_bar(stat = "Identity") +
  labs(x = "Number of Disabled Individuals in Household", y = "Average Level of Food Insecurity", fill = "Average Amount in USD") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

f3 <- ggplot() + geom_boxplot(aes(group = disability, x = disability, y = fexpend, fill = disability), data = cps_fexpend_f) +
  labs(x = "If Disabled Person Lives Within Household", y = "Level of Food Expenditure Within Household", fill = "Disability Level") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_brewer(palette = "BuGn")

grid.arrange(f1, f2, f3, nrow = 1)
```


### Elderly Analysis
```{r, fig.width = 20, fig.height = 12}
d1 <- ggplot(data = cps_fexpend, aes (x = elderly))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "Elderly In Household", y = "Number of Households") +
  scale_fill_brewer(palette = "BuGn")

cps_fexpend_elderly <- cps_fexpend %>% group_by(elderly) %>% summarise(mexp = mean(fexpend))

d2 <- ggplot(aes(x = elderly, y = mexp, fill = mexp), data = cps_fexpend_elderly) + geom_bar(stat = "Identity") + 
  labs(x = "Number of Elderly in Household", y = "Average Amount of Food Expenditure in Household", fill = "Average Amount in USD") + 
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

d3 <- ggplot() + geom_boxplot(aes(group = elderly, x = elderly, y = fexpend, fill = elderly), data = cps_fexpend) +
  labs(x = "Number of Elderly in Household", y = "Amount of Food Expenditure in Household - log scale", fill = "# of Elderly") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_y_log10(labels = scales::dollar_format()) + 
  scale_fill_distiller(palette = "BuGn")

grid.arrange(d1, d2, d3, nrow = 1)
```

### Education Analysis
```{r, fig.width = 20, fig.height = 12}
z1 <- ggplot(data = cps_fexpend, aes(x = education))+ geom_bar(aes(fill = education)) + geom_text(stat = 'count',aes(label=..count..), vjust = -1) +
  labs(x = "Number of Educated Individuals Within Household", y = "Number of Households") +
  scale_fill_distiller(palette = "BuGn")

cps_fexpend_education <- cps_fexpend %>% group_by(education) %>% summarise(med = mean(fexpend))

z2 <- ggplot(aes(x = education, y = med, fill = med), data = cps_fexpend_education) + geom_bar(stat = "Identity") +
  labs(x = "Number of Educated in Household", y = "Average Amount of Food Expenditure in Household", fill = "Average Food Expense USD") +
  scale_y_continuous(labels=scales::dollar_format()) +  scale_fill_distiller(palette = "BuGn", labels=scales::dollar_format())

z3 <- ggplot() + geom_boxplot(aes(group = education, x = education, y = fexpend, fill = education), data = cps_fexpend) +
  scale_y_log10(labels = scales::dollar_format()) + labs(x = "Number of Educated in Household", y = "Average Amount of Food Expenditure in Househod - log scale", fill = "Number of Educated") + scale_fill_distiller(palette = "BuGn")

grid.arrange(z1, z2, z3, nrow = 1)
```


### Employed Analysis
```{r, fig.width = 20, fig.height = 12}
j1 <- ggplot(data = cps_fexpend, aes(x = employed))+ geom_bar() + geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of employed Individuals Within Household", y = "Number of Households")

cps_fexpend_employed <- cps_fexpend %>% group_by(employed) %>% summarise( Average = mean(fexpend))

j2 <- ggplot(aes(x = employed, y = Average, fill = Average), data = cps_fexpend_employed) + geom_bar(stat = "Identity") +
  labs(x = "Number of Employed in Household", y = "Average Amount of Food Expenditure Per Household", fill = "Average Amount in USD") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

j3 <- ggplot() + geom_boxplot(aes(group = employed, x = employed, y = fexpend, fill = employed), data = cps_fexpend) + 
  labs(x = "Number of Employed Within Household - log scale", y = "Food Expenditure Per Household", fill = "Number of Employed") +
  scale_y_log10(labels = scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

grid.arrange(j1, j2, j3, nrow = 1)
```


### Household Size Analysis
```{r, fig.width = 20, fig.height = 12}
s1 <- ggplot(data = cps_fsecurity, aes(x = hhsize)) + geom_histogram(binwidth = 1) + 
  #geom_text(stat = 'count', aes(label = ..count..), vjust = -1) +
  labs(x = "Number of Family Members Household", y = "Number of Households")

cps_fexpend_employed <- cps_fexpend_f %>% group_by(hhsize_f) %>% summarise(Average = mean(fexpend))

s2 <- ggplot(aes(x = hhsize_f, y = Average, fill = Average), data = cps_fexpend_employed) + geom_bar(stat = "Identity") +
  labs(x = "Number of Individuals Within Household", y = "Average Level of Food Insecurity") +
  scale_y_continuous(labels=scales::dollar_format()) + scale_fill_distiller(palette = "BuGn")

s3 <- ggplot() + geom_boxplot(aes(group = round(hhsize,0), x = round(hhsize,0), y = fexpend, fill = round(hhsize,0)), data = cps_fexpend) +
  labs(x = "Number of Individuals Within Household", y = "Food Expenditure Per Household", fill = "Number of Individuals") +
  scale_fill_distiller(palette = "BuGn") + scale_y_continuous(labels=scales::dollar_format())


s4 <- ggplot(data = cps_fexpend, aes(x = hhsize, y = fexpend))+geom_jitter()+ labs(x = "Number of Family Members Within Household", y = "Food Expense" ) +
  scale_y_continuous(labels=scales::dollar_format())

grid.arrange(s1, s2, s3, s4, nrow = 2)
```


Predictive Model - Food Security {data-navmenu="Predictive Models"}
===

Row
--------------
### Desciption:
<font size = "3">
This plot was created from the randomforest for food security, and it lists the variables in their importance in food security determinance.
</font>

Row
--------------
```{r, fig.width = 15, fig.height = 12}
final_forest <- readRDS("final_forest.RDS")

varImpPlot(final_forest, type = 1)
```

Predictive Model - Food Expenditure {data-navmenu="Predictive Models"}
===

Row
--------------
### Desciption:
<font size = "3">
This plot was created from the randomforest for food expenditure per individual for one week, and it lists the variables in their importance in food expenditure determinance.
</font>

Row
--------------
```{r, fig.width = 15, fig.height = 12}
fexpend_final_forest <- readRDS("fexpend_final_forest.RDS")

varImpPlot(fexpend_final_forest, type = 1)
```


Descriptive Model - Food Security {data-navmenu="Descriptive Models"}
=============================================
Row{data-height=200}
---------------------------------------------
### Description:
<font size="3">
This model will not be used for the final prediction, however, it will be used as a representation of what the relationships between
the response variable and explanatory variables are. A Zero Inflation Poisson Model is a regular poisson model except that it accounts for a significant amount of zeroes within the response variable. This is necessary because a majority of houses within the U.S. will have stated that they are not food insecure and therefore be a zero.</font>

Column {.tabset}
---------------------------------------------
### Zero Inflation Poisson Model
<font size="3"> 
Summary 
</font> 
```{r, attr.output='style="max-height: 400px;"'}
fsecurity.glm2 = zeroinfl(fsecurity ~ disability + education + elderly + urban_c | disability + education + elderly + urban_c, data = cps_fsecurity)

summary(fsecurity.glm2)
```

### Exponentiated Coefficients and Confidence Intervals
```{r, attr.output='style="max-height: 200px;"'}
exp(coef(fsecurity.glm2))
```

### Confidence Intervals
```{r, attr.output='style="max-height: 200px;"'}
exp(confint(fsecurity.glm2))
```

### Interpretations for Count Model
<font size = "3">
Disability Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the level of food insecurity increases by a factor of exp(.2593606) = 1.2961011 if there is a disabled person living within the house. This means that if there is a disabled person living within the house, food insecurity of the house increases from the mean by about 30%.

Education Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the mean amount of food insecurity decreases by a factor of exp(-.1114594) = .8945277 for every educated person within the house. This means that for every new educated individual living within the house, food insecurity of the house decreases from the mean by about 10.55%.

Elderly Interpretation (Count Model): With all other variables held constant, for vulnerable individuals, the level of food insecurity decreases by a factor of exp(-.1748214) = .8396069 for every educated person within the house. This means that for every new elderly person living within the house, food insecurity of the house decreases from the mean by about 16.04%. </font>


### Interpretations for Zero-Inflation Model
<font size = "3">
Disability Interpretation (Zero-Inflation Model): With all other variables held constant, the odds a house is food secure is exp(-1.0453348) = 0.3515741. This means that if there is a disabled person living within the household, the odds of a house having a food insecurity level of 0 or greater increases by 64.84%.

Education Interpretation (Zero-Inflation Model): With all other variables held constant, the odds that a household is food secure is exp(0.7795625) = 2.1805181. This means that for every person within the house with an associates degree or higher, the odds of a house having a food insecurity level of 0 or greater decreases by 118%.

Elderly Interpretation (Zero-Inflation Model): With all other variables held constant, the odds that a house is food secure is exp(0.6967225) = 2.0071634. This means that for every elderly person within the house, the odds of a house having a food insecurity level of 0 or greater decreases by 100%.
</font>

Descriptive Model - Food Expenditure {data-navmenu="Descriptive Models"}
==============

Row{data-height=200}
--------------
### Description:
<font size = "3">
This model will not be used for the final prediction, however, it will be used as a representation of what the relationships between
the response variable and explanatory variables are. A Zero Adjusted Gamma Model is a regular gamma model except that it accounts for a significant amount of zeroes within the response variable. This is necessary because a majority of houses within the U.S. have said that they have spent $0 this week per person within the household. Suspected reasons for this would either be people who are getting food from food banks, families that buy food in bulk, or they could be dependants, although this is unlikely since it is an average of a block. 
</font>

Column {.tabset}
-------------
### Zero Adjusted Gamma Model
<font size="3"> 
Summary 
</font> 
```{r, attr.output='style="max-height: 200px;"'}
cps_fexpend$fexpend_0 <- ifelse(cps_fexpend$fexpend == 0, 1, 0)

cps_fexpend_new <- subset(cps_fexpend, fexpend != 0)

fexpend.glm2 <- glm(fexpend ~ hhsize + elderly + employed + disability + education + urban_c, data = cps_fexpend_new, family = Gamma(link = "log"))
fexpend.glm3 <- glm(fexpend_0 ~ hhsize + elderly + employed + disability + education + urban_c, data = cps_fexpend, family = binomial(link = "logit"))

summary(fexpend.glm2)

summary(fexpend.glm3)
```


### Exponential Coefficients
``` {r, attr.output='sytle="max-height: 200px;"'}
exp(coef(fexpend.glm2))
exp(coef(fexpend.glm3))
```

### Confidence Intervals 
```{r, attr.output='sytle="max-height: 200px;"'}
exp(confint(fexpend.glm2))
exp(confint(fexpend.glm3))
```

### Interpretations for Count Model
<font size = "3">
HHSIZE INTERPRETATION (Count Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(-.22794) = .7961747. This means that for every new person that lives within the house the amount of money used to  buy food for each individual decreases by about 20.4% per person compared to the mean.

ELDERLY INTERPRETATION (Count Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(-.04794) = .9531918. This means that for every new elderly person within the household, the amount of money in U.S. dollars used to buy food for each individual decreases by about 4.68% per person compared to the mean.

EMPLOYED INTERPRETATION (Count Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(.06672) = 1.0690010. This means that for every new employed individual within the household, the amount of money in U.S. dollars used to buy food for each individual increase by about 6.9% per person compared to the mean.

DISABILITY INTERPRETATION (Count Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(-0.06584) = .9362819. This means that for every new employed individual within the household, the amount of money in U.S. dollars used to buy food for each individual decreases by about 6.37% per person compared to the mean.

EDUCATION INTERPRETATION (Count Model): With all other variables held constant, for individuals who purchase food, the mean amount of food expenditure changes by a factor of exp(0.12464) = 1.1327362.This means that for every new educated individual within the household, the amount of money in U.S. dollars used to buy food for each individual increases by about 13.27% per person compared to the mean. </font>

### Interpretations for Zero Inflation Model
<font size = "3">
HHSIZE INTERPRETATION (Zero-Inflation Model(Binary)): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.07890) = .9241349. This means that for every new person within the household, the odds of said household spending $0 on food per person decreases by about 7.6%.

ELDERLY INTERPRETATION (Zero-Inflation Model(Binary)): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.01478) = .9853277. This means that for every new elderly person within the household, the odds of said household spending $0 on food per person decreases by about 1.47%. 

 EMPLOYED INTERPRETATION (Zero-Inflation Model(Binary)): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.30615)= .7362754. This means that for every new employed person within the household, the odds of said household spending $0 on food per person decreases by about 26.37%.
 
DISABILITY INTERPRETATION (Zero-Inflation Model(Binary)): With all other variables held constant, the odds of spending $0 changes by a factor of exp(0.25147) = 1.2859122. This means that for every new disabled person within the household, the odds of said household spending $0 on food per person increases by about 28.59%.

EDUCATION INTERPRETATION (Zero-Inflation Model(Binary)): With all other variables held constant, the odds of spending $0 changes by a factor of exp(-0.25484) = .7750404. This means that for every new educated person within the household, the odds of said household spending $0 on food per person decreases by about 22.50%.
</font>

Food Security{data-navmenu="Heat Map Of Predictions"}
===

Inputs {.sidebar}
-----------------------------------------------------------------------
```{r, echo = FALSE}
final_fexpend <- read.csv( "Ryan_Data/final_acs_fexpend.csv")

final_fsecurity <- read.csv("Ryan_Data/final_acs_fsecurity.csv")

State_stuff = data.frame(state.abb, state.name)

final_fsecurity$GEOID = as.character(paste0(final_fsecurity$GEOID, substr(final_fsecurity$location, 13, 13)))

final_fexpend$GEOID = as.character(paste0(final_fexpend$GEOID, substr(final_fexpend$location, 13, 13)))

final_combined <- left_join(final_fexpend, final_fsecurity, by = "GEOID")

final_combined$GEOID <- ifelse(nchar(final_combined$GEOID) == 12, final_combined$GEOID, paste0("0",final_combined$GEOID))

selectInput("State", label = "Select State:", choices = unique(State_stuff$state.name)[order(unique(State_stuff$state.name))], selected = NULL)

```


Row
----
```{r, echo = FALSE}

shp_file <- reactive({ block_groups(state =  State_stuff$state.abb[State_stuff$state.name == input$State]) %>%
    left_join(final_combined, by="GEOID" ) %>%
    rmapshaper::ms_simplify( keep = 0.01, keep_shapes = TRUE) %>%
    #     mutate(COUNTYNAME = convert_fips_to_names(COUNTYFP, states = rep("State", nrow(shp_join)), geo_header="COUNTY"))
    mutate(labels = paste0(
      "<strong> County: </strong> ",
      GEOID, "<br/> ",
      "<strong> Food Expenditure: </strong> $",
      fexpend_predictions, "<br/> ",
      "<strong> Food Insecurity: </strong> ",
      fsecurity_predictions, "<br/> "
    ) %>%
      lapply(htmltools::HTML))
})


#for color scale
limit <- c(min(final_combined$fexpend_predictions,na.rm=TRUE), max(final_combined$fexpend_predictions,na.rm=TRUE))
limit_bin <- c(min(final_combined$fsecurity_predictions,na.rm=TRUE), max(final_combined$fsecurity_predictions,na.rm=TRUE))
pal_bin <- colorBin(
  palette = "RdBu", domain = final_combined$fsecurity_predictions,
  reverse = FALSE,
  bins = seq(0, max(final_combined$fsecurity_predictions, na.rm = TRUE) + .01, by = .20)
)

renderLeaflet({
  leaflet(shp_file(),height=500, width=1000) %>%
    addTiles() %>%
    #setView(lng = 0, lat = 30, zoom = 2) %>%
    addPolygons(
      fillColor = ~ pal_bin(fsecurity_predictions),
      stroke=FALSE,
      color = "white",
      fillOpacity = 0.6,
      label = ~labels,
      highlight = highlightOptions(
        color = "black",
        bringToFront = TRUE
      )
    )%>%
    leaflet::addLegend(
      pal = pal_bin, values = ~fsecurity_predictions,
      opacity = 0.7, title = "Average Level of\nFood Insecurity per\nHousehold",
      position = "topright"
    )
})

```

Row
--------------------

### Predictions
```{r, echo = FALSE, attr.output='style="max-height: 200px;"'}
temp1 <- reactive({ shp_file() %>% as.data.frame%>%dplyr::select(c("GEOID","location.y","fsecurity_predictions", "female.y", "hispanic.y",
                                  "black.y", "kids.y", "elderly.y", "education.y", "employed.y",
                                  "married.y", "disability.y", "urban_c.y"))%>%
                  rename(`Predicted Average` = fsecurity_predictions)

})
renderDataTable({temp1()})
```

Food Expenditure{data-navmenu="Heat Map Of Predictions"}
===

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r, echo = FALSE}
selectInput("State2", label = "Select State:", choices = unique(State_stuff$state.name)[order(unique(State_stuff$state.name))], selected = NULL)
```


Row
----

```{r, echo = FALSE}

shp_file2 <- reactive({ block_groups(state = State_stuff$state.abb[State_stuff$state.name == input$State2]) %>%
    left_join(final_combined, by="GEOID" ) %>%
    rmapshaper::ms_simplify( keep = 0.01, keep_shapes = TRUE) %>%
    #     mutate(COUNTYNAME = convert_fips_to_names(COUNTYFP, states = rep("State", nrow(shp_join)), geo_header="COUNTY"))
    mutate(labels = paste0(
      "<strong> County: </strong> ",
      GEOID, "<br/> ",
      "<strong> Food Expenditure: </strong> $",
      fexpend_predictions, "<br/> ",
      "<strong> Food Insecurity: </strong> ",
      fsecurity_predictions, "<br/> "
    ) %>%
      lapply(htmltools::HTML))
})


#for color scale
limit <- c(min(final_combined$fexpend_predictions,na.rm=TRUE), max(final_combined$fexpend_predictions,na.rm=TRUE))
limit_bin <- c(min(final_combined$fsecurity_predictions,na.rm=TRUE), max(final_combined$fsecurity_predictions,na.rm=TRUE))
pal_bin <- colorBin(
  palette = "RdBu", domain = final_combined$fsecurity_predictions,
  bins = seq(0, max(final_combined$fsecurity_predictions, na.rm = TRUE) + .01, by = .20)
)
log_scale = seq(min(log(final_combined$fexpend_predictions), na.rm = TRUE), max(log(final_combined$fexpend_predictions), na.rm = TRUE), length.out = 11)
pal <- colorBin(
  palette = "RdBu", domain = final_combined$fexpend_predictions,
  reverse=FALSE,
  bins = exp(log_scale)
)

renderLeaflet({
  leaflet(shp_file2(),height=500, width=1000) %>%
    addTiles() %>%
    #setView(lng = 0, lat = 30, zoom = 2) %>%
    addPolygons(
      fillColor = ~ pal(fexpend_predictions),
      stroke=FALSE,
      color = "white",
      fillOpacity = 0.6,
      label = ~labels,
      highlight = highlightOptions(
        color = "black",
        bringToFront = TRUE
      )
    )%>%
    leaflet::addLegend(
      pal = pal, values = ~fexpend_predictions,
      opacity = 0.7, title = "Weekly Average\nFood Expenditure\nPer Person",
      position = "topright"
    )
})


```

Row
--------------------

### Predictions
```{r, echo = FALSE}
temp2 <- reactive({shp_file2() %>%as.data.frame%>% dplyr::select(c("GEOID","location.y", "fexpend_predictions","female.y", "hhsize.y", "hispanic.y","black.y", "kids.y", "elderly.y", "education.y", "employed.y","married.y", "disability.y", "urban_c.y"))%>%
                  rename(`Predicted Expenditure` = fexpend_predictions)
})

renderDataTable({temp2()})
```