---
title: "Food Dashboard"
output: 
  flexdashboard::flex_dashboard:
    runtime: shiny
  orientation: rows
---
```{r setup, include=FALSE}
#multiple pages within one dashboard
#https://rmarkdown.rstudio.com/flexdashboard/layouts.html#multiple_pages
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(reshape2)
#NOTE: "-----------------------------------------------------------------------" is the section delimiter
#NOTE: "===================================== " is the PAGE delimiter

```
Data Overview
===

Row {data-height=200}
-----------------------------------------------------------------------

### Goal of Research:

The goal of this analysis is to find determinants of food insecurity. The reason for doing so is humanitarian in nature as the
increase in efficacy of food distribution centers by determining higher risk areas should decrease the level of food 


### 



Important Variables
===


Row {data-height=400}
---


### Disability Analysis


```{r, fig.width = 15}
ggplot(data = cps_fsecurity, aes(x = disability_cat, fill = disability_cat))+geom_bar() + geom_text(stat = 'count',aes(label=..count..), vjust = -1) + 
  labs(x = "Disabled Individual Living Within Household", y = "Number of Households", fill = "If Disabled") + scale_fill_brewer(palette = "Blues")
```



Row {data-height=400}


Predictive Model
===





Descriptive Model
===




Heat Map & Prediction
===






