---
title: "First_Progress_Report"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- asdfasdfasdf
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```




# Start sthe slides here and use the top stuff as a reference on how to proceed 

ggplot(data = cps_fsecurity, aes(x = urban_c)) + 
  geom_bar() + 
  geom_text(stat = 'count', aes(label = ..count..), hjust = -1) +
  labs(x = "", y =  "Count") +#LRF: add labels 
  coord_flip() #the x axis is smooshed - thijs helps with that

#LRF: take a look at this - don't have to create all those temporary datasets
cps_fsecurity %>% 
  group_by(urban_c) %>% 
  summarise(Average = mean(fsecurity)) %>%#LRF: pipe the data right into ggplot!
ggplot(aes(x = urban_c, y = Average, fill = Average)) +
  geom_bar(stat = "Identity")  +
  coord_flip()+ #LRF: again, the x axis was smooshed
  labs(x = "", y = "Proportion of Food Insecure")#LRF: axis labels!

  
ggplot(data = cps_fexpend, aes(x = urban_c)) + 
  geom_bar() + 
  geom_text(stat = 'count', aes(label = ..count..), hjust = -1) +
  coord_flip()+
  labs(x = "", y = "Count")

#SEE HOW I'VE CHANGED THIS
cps_fexpend  %>% 
  group_by(urban_c) %>% 
  summarise(Average = mean(fexpend))%>% #LRF piping data right into plot
ggplot(aes(x = urban_c, y = Average, fill = Average)) + 
  geom_bar(stat = "Identity") +
  coord_flip()+ #LRF
  scale_y_continuous(labels=scales::dollar_format())+#LRF: GOOGLE "ggplot2 scale axis dollar"
labs(x = "", y = "Average Spent Per Week, Person") + #gLRF: ood labels
  theme(legend.position = "none") #LRF: legend not necessary here
#  labs(x = "Number of Individuals Within Household", y = "Average Level of Food Insecurity")
